### 2014-06-01 ACL2014/system.tex
section 3 edits
updating problem statement section

\section{Problem Statement}
We start with a corpus of user queries being represented as vectors
over some domain (e.g. URLs); similarity between queries is to be measured using cosine between the corresponding
vectors. We are also given a set of queries $Q$. 
We are interested in developing a batch process to return a \emph{small}
set of candidate neighbors for each query $q \in Q$ such that the following
are satisfied: 1) the similarity of any returned candidate to the query $q$ is larger than a user-specified 
similarity threshold $\tau$; 2) the set of candidates returned has a large recall. 
Note that, although we return an approximate set of near neighbors, our requirement forces us to 
return a candidate set that has $100\%$ precision. Our aim is to achieve a large recall, while using
a scalable efficient algorithm.

The exact brute force algorithm to solve the above problem would be to compute
similarities between each $q\in Q$ and all queries in the corpus, and return 
all pairs that have similarities higher than the threshold $\tau$. Computing similarities
for all pairs is computationally infeasible on a single computer even if the size of $Q$ is of the order of few thousands
and the corpus size is hundreds of millions. Even in a distributed setting such as Hadoop, the 
resulting communication needed between machines makes this strategy impractical.

Our aim is to empirically study a set of locality sensitive hashing techniques
that enable us to return a set of candidate neighbors while performing
a much smaller (theoretically \emph{sublinear}) set of comparisons. 
In order to tackle this scalability problem, we explore the 
combination of distributed computation using a map-reduce platform (Hadoop) as well as
locality sensitive hashing (\lsh) algorithms. 
We explore a few commonly known variants of \lsh and suggest a few novel variants
that are suitable to the map-reduce platform. 
The methods that we propose meets the practical requirements of a real life 
search engine backend, and demonstrates how to use locality sensitive hashing
on a distributed platform.

@@

\subsection{Reusing Hash Functions}
\label{subsec:resuseHash}
In Section \ref{subsec:vlsh:cosine}, we showed that vanilla \lsh requires $L\times K$ hash functions. 
However generating hash functions is computationally expensive as 
it takes time to read all features and 
evaluate hash functions over all those features to generate a single bit. 
To minimize the number of hash functions computations, 
we use a trick from Andoni and Indyk \shortcite{Andoni08CACM} 
in which hash functions are reused to generate $L$ tables. $K$ is assumed to be even and $R \approx	\sqrt L$. 
We generate $f(q)=(h_1(q),h_2(q),\dots,h_{K/2}(q)))$ of length $k/2$. 
Next, we define $g(q)=(f_a,f_b)$, where $1\leq a < b \leq R$. Using such pairing, we
can thus generate $L= \frac{R (R -1)}{2}$ hash indices. 
This scheme requires O($K  \sqrt L$) hash functions, instead of O($KL$).    
For rest of this paper, we use the above trick to generate $L$ hash tables with bucket indices of width $K$ bits.  
\begin{table}
\centering
{
\small \addtolength{\tabcolsep}{-4.5pt}
\begin{tabular}{ll}
\hline
\hline
 \textbf{Symbol} &  \textbf{Description} \\
\hline 
$N$ & \# of query terms \\
$D$ & \# of features i.e. all clicked unique urls \\
\hline
\multirow{3}{*}{$K$} & \# of hash functions concatenated together \\ 
   & $g(q)=(h_1(q),h_2(q),\dots,h_k(q)))$ \\ 
	 & to generate the index of a table \\
\hline
 \multirow{2}{*}{$L$} & \# of tables generated independently \\ 
	  &  with $g_j(q)$ index, $\forall 1 \leq j \leq  L$  \\
\hline
$F$ & \# of bits flipped, $\forall 1 \leq j \leq  L$   \\
$\tau$ & $\tau$ threshold \\
Recall & fraction of similar candidates retrieved \\
Comparisons & Avg \# of pairwise comparisons per query  \\
\hline 
\end{tabular}
\caption{\footnotesize{Major Notations}}

\label{tab:notation}
\end{table}

>>>>

\newcommand{\corpus}{$C$\xspace}

\section{Problem Statement}
We start with user query logs \corpus having query vectors collected from 
a commercial search engines over some domain (e.g. URLs); similarity between queries is to be measured using cosine between the corresponding
vectors. 
The problem we formulate here is given a set of queries $Q$ and similarity threshold $\tau$,
we are interested in developing a batch process to return a \emph{small}
set $T$ of candidate neighbors from \corpus for each query $q \in Q$ such that the followings
are satisfied: 1) $T = \{ l \mid s(l, q) \geq \tau,  l \in C  \}$, where  $s(q_1, q_2)$ is a function to
compute a similarity score between query $q_1$ and $q_2$
; 2) $T$ achieves $100\%$ precision with ``large'' recall. 
That is, our aim is to achieve a large recall, while using
a scalable efficient algorithm.

The exact brute force algorithm to solve the above problem would be to compute
similarities between each $q\in Q$ and all queries in \corpus, and return 
all pairs that have similarities higher than the threshold $\tau$. Computing similarities
for all pairs is computationally infeasible on a single computer even if the size of $Q$ is of the order of few thousands
and the size of \corpus is hundreds of millions. Even in a distributed setting such as Hadoop, the 
resulting communication needed between machines makes this strategy impractical.  This in turn highlights ``linear'' computational complexity is not always acceptable to a certain problem domains.

Our aim is to empirically study a set of locality sensitive hashing techniques
that enable us to return a set of candidate neighbors while performing
a much smaller (theoretically \emph{sublinear}) set of comparisons. 
In order to tackle this scalability problem, we explore the 
combination of distributed computation using a map-reduce platform (Hadoop) as well as
locality sensitive hashing (\lsh) algorithms. 
We explore a few commonly known variants of \lsh and suggest several novel variants
that are suitable to the map-reduce platform. 
The methods that we propose meets the practical requirements of a real life 
search engine backend, and demonstrates how to use locality sensitive hashing
on a distributed platform.

@@

\subsection{Reusing Hash Functions}
\label{subsec:resuseHash}
In Section \ref{subsec:vlsh:cosine}, we showed that vanilla \lsh requires $L\times K$ hash functions. 
However generating hash functions is computationally expensive as 
it takes time to read all features and 
evaluate hash functions over all those features to generate a single bit. 
To minimize the number of hash functions computations, 
we use a trick from Andoni and Indyk \shortcite{Andoni08CACM} 
in which hash functions are reused to generate $L$ tables. $K$ is assumed to be even and $R \approx	\sqrt L$. 
We generate $f(q)=(h_1(q),h_2(q),\dots,h_{K/2}(q)))$ of length $k/2$. 
Next, we define $g(q)=(f_a,f_b)$, where $1\leq a < b \leq R$. Using such pairing, we
can thus generate $L= \frac{R (R -1)}{2}$ hash indices. 
This scheme requires O($K  \sqrt L$) hash functions, instead of O($KL$).    
For rest of this paper, we use the above trick to generate $L$ hash tables with bucket indices of width $K$ bits.  
\begin{table}
\centering
{
\small \addtolength{\tabcolsep}{-4.5pt}
\begin{tabular}{ll}
\hline
\hline
 \textbf{Symbol} &  \textbf{Description} \\
\hline 
$N$ & \# of query terms \\
$D$ & \# of features i.e. all clicked unique urls \\
\hline
\multirow{3}{*}{$K$} & \# of hash functions concatenated together \\ 
   & $g(q)=(h_1(q),h_2(q),\dots,h_k(q)))$ \\ 
	 & to generate the index of a table \\
\hline
 \multirow{2}{*}{$L$} & \# of tables generated independently \\ 
	  &  with $g_j(q)$ index, $\forall 1 \leq j \leq  L$  \\
\hline
$F$ & \# of bits flipped, $\forall 1 \leq j \leq  L$   \\
$\tau$ & $\tau$ threshold \\
Recall & fraction of similar candidates retrieved \\
Comparisons & Avg \# of pairwise comparisons per query  \\
\hline 
\end{tabular}
\caption{\footnotesize{Major Notations}}
}
\label{tab:notation}
\end{table}
